\chapter*{Acknowledgements}
% \thispagestyle{empty}

Completing this dissertation would not have been possible without the support of many kind, driven, and supportive researchers, to whom I want express my gratitude. First, I want to thank to my advisor, Zhi Ding, who secured the funding for my degree\footnote{The research conducted in this dissertation was supported by the National Science Foundation under the following grants: ECCS-1711823, ECCS-2029027, and CNS-2002937.} and who always challenged me to be a better researcher. I would also like to thank Lifeng Lai and Khaled Abdel-Ghaffar, who served on both my qualifying exam and dissertation committees, as well as Daniela Cabric and Zhaojun Bai, who graciously served as members of my qualifying exam committee. Finally, I thank my lab mates and collaborators, particularly Zhenyu Liu, Yu-Chien Lin, Carlos Feres, Siyu Qi, and Lahiru Chamain, who provided support, feedback, and fruitful discussion which helped shaped the research described in this dissertation.

Outside of my advisors and colleagues in research, I had numerous other mentors and collaborators during my time at UC Davis. I want to thank Hooman Rashtian, with whom I taught as a teaching assistant for several undergraduate courses and the COSMOS program. Hooman's passion as an educator helped me develop an appreciation of engineering education, and I enjoyed working with him very much. I would also like to acknowledge my colleagues in 2021-22 cohort of the Professors for the Future Fellowship as well as the organizers of the program, Teresa Dillinger, Ellen Hartigan-O'Connor, and David Blancha. Finally, I would like to thank my colleagues and consultees from the Graduate Writing Fellows program as well as the head of the program, Katherine Gossett.

Finally, I want to thank my family and friends who have tirelessly supported me on my doctoral journey. I want to start by thanking my parents, Michael and Gail, who have provided me with countless opportunities to succeed. I also want to thank my brother, Zach, who inspired me to embark on this doctoral journey, who always advocated for me, offered me advice, and continues to inspire me to this day. I will also thank my friends from Olin College, Ankur, Carly, Brandon, Abe, and Wanyi (an honorary Oliner), my best friend from Carnegie Mellon, Josh, and my friends from Temple University Japan, Ty and Casey. Most importantly, I thank my partner, Misha Nguyen, who keeps me grounded and who fills my life with love and light.

% \clearpage
