\newpage
\centerline{\Huge \bf {Acknowledgments}} \vspace{18pt}
\addcontentsline{toc}{section}{Acknowledgments}


This dissertation along with the work invested into it would not have been possible without the kind support of many people.

First of all, I would like to express my most sincere gratitude to my advisor,
Prof. Zhi Ding, 
for his continuous guidance and encouragement throughout my Ph.D. studies and research, for his patience, motivation, inspiration and endless support.
I am truly thankful for all his help in selecting research topics, suggesting technical insights, proofreading manuscripts, offering teaching opportunities, providing financial support, and overall, for his immense knowledge and wisdom. 
From him I have not only learned an effective approach for academic research, but also how to engage with others while doing so, in an encouraging and profoundly motivating manner.
He is the best mentor I could ever have in this endeavor, and I deeply appreciate all these years under his tutelage.

I am also indebted to the members of my dissertation committee, Prof. Bernard Levy and Prof. Khaled Abdel-Ghaffar, for the time they spent in evaluating my dissertation. 
I have taken classes with both of them, which sparked my interest and allowed me to dive deeper in my research. 
I am also thankful to Prof. Lifeng Lai and Prof. Xin Liu for their service on my qualifying exam. 

I would also like to thank all my friends and colleagues at the Broadband Radio Access Technology Lab (BRAT-Lab), especially (in no particular order), 
Wenhao Wu, 
Chen Jiang,
Kun Wang, 
Taha Bouchoucha, 
Mohammadamin Farajzadehjalali, 
Hesham Mostafa Elsayed Elmaghraby, 
Xintong Ling,
Pei-Rong Li, 
Songyang Zhang, 
Qinwen Deng, 
Lahiru D. Chamain Hewa Gamage, 
Siyu Qi, 
Mason del Rosario, 
GE Yao, 
Zhenyu Liu, 
Hui-Ying Siao, 
Dylan Shadduck, 
Jonathan Tivald, 
Kuangang Fan, 
and Qing Wei, 
for their great support, technical advice and stimulating discussions.

I would also like to acknowledge all my friends in Davis who have supported me throughout the last five years, and made my Ph.D. experience even better. 

None of my accomplishments would have been possible without the
everlasting love, encouragement and support of my family. The caring and
guidance of my beloved parents, Carlos and Carolina, have been the foundation of who I am today, and even after their passing they continue to be my guides. My son Carlos Le\'on has been a shining light during these years.

And above all, I would like to thank my wife Constanza for her constant love and support, and for keeping me sane over the past few years. Thank you for being my best friend. I owe you everything.

The work invested into this dissertation was supported in part by the National
Science Foundation under Grants No. 1711823, No. 1824553, No. 2009001, and No. 2029027. 
My Ph.D. studies and research were also supported by the National Agency for Research and Development (ANID) / Scholarship Program / DOCTORADO BECAS CHILE / 2016 72170648.
